Reaction-Advection-Diffusion Models of Air Pollution

### 2024-04-28 paper.tex
9.5

$$
    \int_0^{L_y} K'_y Y'_m(y)Y_n(y) dy = \eta{mn}
$$

@@

@@

One of the recent developments in the study of air pollution was the creation of computationally efficient models for computing the harm to human health from
diffusion. A prevalent model that has been implemented in industry by companies such as WattTime is InMAP \url{(http://spatialmodel.com/)}. 
InMAP is a model that evaluates pollution by simplifying typical CTMs via using an annual-average basis for its results.

>>>>

$$
    \int_0^{L_y} K'_y Y'_m(y)Y_n(y) dy = \eta_{mn}
$$

@@

This result turned out to be very accurate against experimental data (Buske, 2012). 
The equation shown above was ran against data from Copenhagen, Denmark, and Kinkaid, Illinois. 
The difference between the predicted and actual results turned out to be within an appropriate 
error bound, leading for the authors to conclude that analytical solutions to the advection differential
equation (ADE) may be possible. The potential for creating an accurate and computationally efficient model
for modeling air pollution dispersion encouraged more research into new methods of CTMs.

@@

One of the recent developments in the study of air pollution was the creation of computationally efficient models for computing the harm to human health from
diffusion. A prevalent model that has been implemented in industry by companies such as WattTime is InMAP \url{(http://spatialmodel.com/)}. 
InMAP is a model that evaluates pollution by simplifying typical CTMs via using an annual-average basis for its results. InMAP itself
was based on a revolutionary CTM, WRF-Chem (Grell et al., 2005).

\subsection{How does WRF-Chem work?}

WRF-Chem is a CTM that is based on the Weather Research and Forecasting Model (WRF). The WRF was initially created by
researchers in the late 1990s via a collaboration between the National Center for Atmospheric Research (NCAR), 
the National Oceanic and Atmospheric Administration, and many others. WRF is most useful for its capability to 
apply a mathematical model via a direct simulation with idealized conditions or use actual atmospheric data.
It is used at many meterological companies as well as in solar companies, laboratories, and many other locations.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.3\textwidth]{wrfchem.png}
    \caption{An example of the graphics produced by WRF Chem.}
    \label{fig:atmospheric_box_model}
 \end{figure}

WRF-Chem combines this with the ability to trace $PM_{2.5}$. It is very numerically consistent and its forecasts are 
stronger than many models beforehand. However, like many other CTMs, WRF-Chem is very computationally expensive.

\subsection{InMAP's solutions}

To solve many of the computational issues that WRF-Chem had, InMAP makes several key simplifications. InMAP simplifies many of the 
marginal emissions rates into annual averages, creating an algorithm that can be run much more efficiently. In addition to being more
computationally efficient, InMAP also creates more spatially detailed results. This efficiency is achieved by using a variable-resolution grid:
areas with a higher density of people are modeled with a higher spacial detail than regions of lower density. The algorithm is able to be downloaded
and is programmed using the Go programming language; it is designed to be simple enough for a beginner user to learn. The authors claim that this is a
novel approach to CTMs.

\begin{figure}[h]
    \centering
    \includegraphics[width=\textwidth]{inmap.png}
    \caption{How InMAP splits the US into regions. Note how regions with higher population density are modeled as smaller grid
    regions, and regions with lower density have larger grid regions}
    \label{fig:grid}
 \end{figure}

\subsection{The diffusion equation}

InMAP utilizes a reaction-advection-diffusion equation to find its air pollution concentrations:

\begin{equation}
    \frac{\partial C_i}{\partial t} = \underbrace{\nabla \cdot (D\nabla C_i)}_{\text{Diffusion Term}} - \underbrace{\nabla \cdot (\vec{v}C_i)}_{\text{Advection Term}} + \underbrace{\sum_{j=1}^{n} R_{ij}}_{\text{Reaction Terms}} + \underbrace{E_i}_{\text{Source/Sink Term}} - \underbrace{d_i}_{\text{Decay Term}}
\end{equation}

\begin{itemize}
    \item $C_i$ is the concentration of one of $n$ model pollutant species.
    \item $D$ is a molecular diffusion coefficient (neglected here as a negligible source of chemical transport in the atmosphere compared to advection).
    \item $\vec{v}$ is the wind vector.
    \item $R_{ij}$ is the net formation rate of species $i$ from species $j$.
    \item $E_i$ is pollutant emission.
    \item $d_i$ is pollutant removal via wet and dry deposition.
\end{itemize}

InMAP solves this equation for the concentrations of $\text{PM}_{2.5}$ and many other harmful chemicals emitted. To find the concentrations,
InMAP runs in each grid region at a discrete step interval. During each step, the following are performed:

\begin{enumerate}
    \item InMAP finds the flux that comes from new emissions. This is done using the plume model, as referenced before. 
    \item The algorithm then calculates how changes in pollutant concentrations are affected by physical and chemical processes including advection, turbulent mixing, etc.
    \item Each process listed above uses a snapshot of the concentration before the time step had occurred to avoid combining changes to concentrations at the same time. 
  \end{enumerate}

\subsection{How InMAP takes in data}

InMAP first takes in the data and runs it through a preprocessor that is an extension of WRF-Chem. This reduces the complexity of the operations
that need to be ran after. Since the CTM is computationally expensive to run, the input data time step is once per WRF-Chem simulation hour. 
After this, the InMAP algorithm runs stepwise, adjusting the relevant terms of the equation as needed. One of the more interesting features is the 
methodology of updating the advection wind vectors. The wind speed updates at too fast of a rate for any computer to calculate using discrete steps.
To solve this problem, Reynolds decomposition is used: this is a mathematical technique that uses the expected value
of a wind stream and its deviations to calculate an accurate result. For example, if a variable $x$ is decomposed, it would be 
represented as the following equation:

\begin{equation}
    x = \overline{x} + x'
\end{equation}

\begin{itemize}
    \item $\overline{x}$ is the expected value of $x$
    \item $x'$ is deviations in x.
\end{itemize}

An example of the output of the algorithm is shown below:

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{inmap2.png}
    \caption{The result of the model when running with $\text{PM}_{2.5}$ as the specified pollutant.}
    \label{fig:grid2}
 \end{figure}
